---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup}
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(boundr)
library(tmap)
library(ggplot2)
```

# welcome to boundr

<!-- badges: start -->

<!-- badges: end -->

### Retrieve area boundaries and data from the ONS Open Geography Portal

`boundr` was previously
[`jogger`](https://github.com/francisbarton/jogger).

The main purpose of this package is to download area lookups and
boundaries (in GeoJSON format) using the ONS Open Geography API, for all
areas at a specified level within a specified area.

The main script will return a data frame with the sub-area geometry
column, as an `sf` object ready to be visualised as a map.

### Package logic

The structure of the project looks a bit like this:

    bounds() [main UI function]
       ^
       |
       |
        <-------  create_lookup_table()
       |                ^
       |                |
       |                 <--------- return_query_data()
       |                |
       |                 <--------- return_lookup_query_info() 
       |                                       \
       |                                        \
        <-------  return_bounds_data()           <----- opengeo_schema [data]
                        ^                       /              ^
                        |                      /               |
                         <--------  pull_geo_query_url()       |
                                                               |
                                                         build_schema()

When you call `bounds()` you specify a lower level area (eg ward) and a higher
level area (eg local authority), and you specify either the name of the higher
level area (or areas) or its code.

`return_lookup_query_info()` then finds the API query URL of a suitable lookup
table - one that contains columns for both your lower and higher level areas.
It does this by filtering `opengeo_schema`, which is a cached copy of the
various datasets available from the Open Geography API Services list.
This schema is available as internal data in the package - but may need
updating.

`create_lookup_table()` then builds a lookup table (a tibble) based on all the
areas you have said you are interested in.
At the same time, `return_bounds_data()` will - if you have specified you want
spatial boundaries data for your areas - retrieve the boundary data at your
chosen resolution for your lower level areas.
These will then be joined onto the lookup table and provided to you as an
`sf` tibble. 


## Installation

You can install this package from the `R` console by entering

  `remotes::install_github("francisbarton/boundr")`

if you have the `remotes` package installed.

## Examples

### Basic lookup of areas within a larger area, by name

Returns sfc tibble with latest available data

```{r example-1}
#| cache: true

bounds("msoa", "lad", "Swansea")

```


### Lookup areas with older dates

Trial and error may be required - or start by looking up what is available on the OpenGeography site.

```{r example-2}
#| cache: true

bounds("msoa", "lad", "Shepway", lookup_year = 2011, within_year = 2015)
```


### Alternative ways to achieve a goal

In this case, the first option is notably quicker than the second, because they use different tables to create the lookup table.

The second option fails to return boundaries unless the year option is included, as it creates a lookup table for rgn23 and ctry23 but boundaries are not available for rgn23 (only rgn22).

Different columns are returned by each command.

```{r example-3}
#| cache: true

bounds("rgn", country_filter = "EN", resolution = "BUC")
bounds("rgn", "ctry", "England", lookup_year = 2022, resolution = "BUC")
```


### A lookup table for Welsh local authorities to Senedd electoral regions

Using `return_with = "full"` includes the Welsh language data columns.

```{r example-4}
#| cache: true

create_lookup_table("ua", "sener", return_width = "full") |>
  dplyr::filter(sener22nm == "South Wales West")

```


```{r example-5}
#| cache: true
#| messages: false

bounds("parish", "utla", "Isles of Scilly") |>
  tmap::qtm()
```


```{r example-6}
#| cache: true

bounds("spr")
```



```{r example-7}
#| cache: true
#| messages: false

sb <- bounds("msoa", "lad", "Swindon")
sp <- points("msoa", "utla", "Swindon")

sb |>
  ggplot2::ggplot() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = sp, colour = "orange") +
  ggplot2::theme_void()

```



```{r example-8}
#| cache: true

bounds("npark", within_names = "Bannau Brycheiniog", resolution = "BUC") |>
  ggplot2::ggplot() +
  ggplot2::geom_sf()
```

## Contributing

Suggestions for improvements are welcome, preferably posted as an issue
here on GitHub.
Contributions as pull requests are also welcome.

Please note that `boundr` is released with a [Contributor Code of
Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## News

See [NEWS.md](NEWS.md) for version notes.

### Licences

The code in this repo is MIT licensed.

The data that the code helps you retrieve is issued under a variety of
licences, including:

-   the [Open Government Licence
    v3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)

> Licensing statement [as stipulated by the
> ONS](https://www.ons.gov.uk/methodology/geography/licences):
>
> -   Source: Office for National Statistics licensed under the [Open
>     Government Licence
>     v3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)
> -   Contains OS data © Crown copyright and database right 2021.
